### 2020-05-18 probabilisticProgrammingMartingales.tex
Questions about "partial ranking function is a restriction of a ranking function".

\newpage

@@

    \item $\big(\lambda z. \Theta \; (\tY \, \Theta) \;  z\big) \; (\underline{i} + 1)$.
\end{enumerate}

Even in this simple case, defining a ranking function explicitly is awkward because of the number of cases, although in most cases, because the value need only be greater than or equal to that of the next term in sequence, it suffices to take the ranking function as taking only 3 distinct values.

@@

Define $(\next(N,s),\_) = \red^n(N,s)$ for the least $n \geq 0$ such that it's in the domain of $f_1$, and $g(N,s) = \left | \{m < n \mid \red^M(N,s) \text{ is of the form } (E[\tY N'],s') \} \right |$. $\next$ is well-defined (i.e.~$n$ is finite) for all $N \in Rch(M)$ by induction on the path from $M$ to $N$, by the third condition on partial ranking functions. Define $f_2(N) = \int_S f_1(\next(N,s)) + g(N,s) \, \mu(\mathrm d s)$. The (total) function $f_2$ agrees with $f$ on $f$'s domain, and it is a ranking function on $M$ (in fact, the least ranking function of which $f$ is a restriction).
\end{proof}

@@

>>>>

@@

    \item $\big(\lambda z. \Theta \; (\tY \, \Theta) \;  z\big) \; (\underline{i} + 1)$

    \item $\Big(\lambda z. \big(
    \lambda n: 
    \tif{\tsample - 0.5 < 0}{n}{(\tY \, \Theta) \; (n+1)} 
    \big) \, z \Big)
    \; \underline{i}$
\end{enumerate}

\lo{So we have:
\[
\begin{array}{ll}
1 \to 11 \to 4 \to 5 \to 6 \to 7 \to 8\\
7 \to 9 \to 11 \to \cdots
\end{array}
\]
You have 2, 3 and 10 because your $\tY$-redex rule is
\[
\tY \lambda x. s \to \lambda z. (\lambda x.s)\, (\tY \lambda x. s) \, z
\]
which is not the official definition.}

Even in this simple case, defining a ranking function explicitly is awkward because of the number of cases, although in most cases, because the value need only be greater than or equal to that of the next term in sequence, it suffices to take the ranking function as taking only 3 distinct values.
\lo{This is only true for partial ranking function.}

@@

Define $(\next(N,s),\_) = \red^n(N,s)$ for the least $n \geq 0$ such that it's in the domain of $f_1$, and $g(N,s) = \left | \{m < n \mid \red^m(N,s) \text{ is of the form } (E[\tY N'],s') \} \right |$. 
The function $\next$ is well-defined (i.e.~$n$ is finite) for all $N \in Rch(M)$ by induction on the path from $M$ to $N$, by the third condition on partial ranking functions. Define $f_2(N) = \int_S f_1(\next(N,s)) + g(N,s) \, \mu(\mathrm d s)$. The (total) function $f_2$ agrees with $f$ on $f$'s domain, and it is a ranking function on $M$ (in fact, the least ranking function of which $f$ is a restriction).
\end{proof}

@@

\lo{Call the above partial function $h$.
I assume that $h: (\tY f, n. n \oplus_{0.5} f(n+1)) N \mapsto 2$ for all $N$ (such that $(\tY f, n. n \oplus_{0.5} f(n+1)) N$ is reachable from $M$), 
and the second and third clauses hold for all $i \geq 0$.

Maybe I have misread / misunderstood the preceding, but I don't see why $h$ is a restriction of a ranking function. 
Suppose $h$ is a restriction of $H$.
Consider the reduction sequence (writing $\Xi = (\tY f, n. n \oplus_{0.5} f(n+1))$):
\[
\Xi \, 0 
\xrightarrow{\tY} 
0 \oplus_{0.5} \Xi \, 1 
\to 
\Xi \, 1
\xrightarrow{\tY} 
1 \oplus_{0.5} \Xi \, 2
\to
\Xi \, 2
\xrightarrow{\tY} 
2 \oplus_{0.5} \Xi \, 3
\]
Since $H(\Xi \, 0) = 2$, we must have 
$H(\Xi \, 1) \leq 1$ and $H(\Xi \, 2) \leq 0$, contradicting the restriction assumption.
}
